Measuring contrast processing in the visual system using SSVEP

### 2024-04-13 review.qmd
Added a few bits

To illustrate the effect of contrast gain control [@Heeger1992], we include a second component that contributes to the gain pool of the first. At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).

@@

This possibility was recognised in the early days of the technique [@Regan1988]. Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data. ... MORE HERE - Although this idea was suggested as early as 1988 (probably earlier) not many people seem to have done it much. @Candy2001; @Busse2009; @Tsai2012; @Baker2017 *** SEE ALSO RECENT WORK FROM WINAWER LAB - MUCH OF IT IS 'FAST' VEPs ***

@@

Parkinson's @Afsari2014

```{python}
#| label: fig-plotclinical
#| fig-cap: "Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d) shows data from *Drosophila melanogaster* (fruit flies) in Afsari et al 2014. One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity compared to control animals."
#| fig-cap-location: bottom

>>>>

To illustrate the effect of contrast gain control [@Heeger1992], we include a second component (a 'mask') that contributes to the gain pool of the first (the 'target'). The mask stimulus will suppress the response at the target frequency, reducing its amplitude. The suppression is reciprocal - activity at the mask frequency is also reduced by the presence of the target, in a contrast-dependent manner [@Busse2009]. At a single pair of (matched) contrast levels, we see a complex pattern of intermodulation terms in the FT (@fig-intermodulation).

@@

This possibility was recognised in the early days of the technique [@Regan1988]. Although characterising the complete set of computations along entire processing pathway is challenging, careful parametric variation of the input stimuli does allow us to fit models of early visual gain control using SSVEP data. For example, @Tsai2012 demonstrated that a gain control model gave a good account of the pattern of intermodulation responses produced by two overlaid patterns flickering at different frequencies. This was achieved by passing full stimulus waveforms through the transducer nonlinearity, and calculating the Fourier spectrum of the model output. Our own work on signal combination across eyes and space similarly demonstrated close correspondence between the predictions of a computational model and empirical data in humans [@Baker2017]. More detailed modelling of intracortical recordings [@Groen2022] has revealed details of the timecourse of gain control effects and adaptation, specifically that normalization is delayed slightly relative to the initial visual response.

@@

Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models. @Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d). The authors theorised that differences in early gain control during development might lead to visual deficits later in life. Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms manifest. The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments. SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.

#| label: fig-plotclinical
#| fig-cap: "Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al. (2014). One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black)."
#| fig-cap-location: bottom
